MSc Statistical Rob correction -Jan29

### 2018-02-08 MSc_Statistical_Rob_correction_-Jan29.Rmd
PCA and working on field data

```{r}
mod1<-PlantLength$Length_plant~PlantLength$Location+(1|PlantLength$Species)+(1|PlantLength$Population)
row.has.na <- apply(PlantLength, 1, function(x){any(is.na(x))})
sum(row.has.na) 
PlantLength.filtered <- PlantLength[!row.has.na,]# Removed all the rows with NAs

mod1<-lmer(log(PlantLength.filtered$Length_plant)~PlantLength.filtered$Location+(1|PlantLength.filtered$Species)+(1|PlantLength.filtered$Population))
mod2<-lmer(log(PlantLength.filtered$Length_plant)~PlantLength.filtered$Location+(1|PlantLength.filtered$Species))
mod3<-lmer(log(PlantLength.filtered$Length_plant)~1+(1|PlantLength.filtered$Species)+(1|PlantLength.filtered$Population))
# Made these three models, one is all of them, one no population, and one no in/ out
summary(mod1)
anova(mod1,mod2) # One ANOVA with model 1, one with model 2 to compare if they are statistically significantly different from each other
qplot(predict(mod1),log(PlantLength.filtered$Length_plant)) #is this model normal? NO
qplot(predict(mod1),log(PlantLength.filtered$Length_plant)-predict(mod1)) # Make this model normal by log() the y variable.
cor(predict(mod1),PlantLength.filtered$Length_plant)^2 # Are the correlation coefficients meaningful?
cor(predict(mod2),PlantLength.filtered$Length_plant)^2 # Are the correlation coefficients meaningful wihout pops?

anova(mod1,mod3) # Repeat with a model without the in/ out component.
cor(predict(mod3),PlantLength.filtered$Length_plant)^2 # how much different does it make when we consider GM?
```

@@

```{r}
colnames(FloristicSurvey)[1] <- "Population" # population loads in weirdly, I am just arranging it to not have quotes around. so odd
FloristicSurvey$Population<-as.factor(FloristicSurvey$Population) # Make the population a factor and not an interger
FloristicSurvey[is.na(FloristicSurvey)] <- 0
sum(row.has.na <- apply(FloristicSurvey, 1, function(x){any(is.na(x))}))

@@

### Linear discriminate analysis
```{r}
# Discriminant function analysis is used (general LDA format) because i have a cat. dependent and a binary independent. (it needs binary or continuous!)
# you might have to bin covers because it only works with categorical, so 0-5% none, 6-40 low etc.
FloristicSurvey$GM_Coverage_category<-cut(FloristicSurvey$GM_Coverage, c(-Inf,20,50,80,100), labels = c("None", "Med_Low", "Med_High", "Overtaken"))
table(FloristicSurvey$GM_Coverage_category)
# [0-20] None
# ]20-50] Medium low
# ]50-80] Medium high
# ]80-100] Overtaken
```

### LDA
```{r}
#LDA
FloristicSurveysmall<-data.frame(FloristicSurvey[,4:56])
# Linear discriminant function analysis

@@

### Cluster analysis
```{r}
# Ward Hierarchical Clustering with Bootstrapped p values
#do they cluster by population?
FloristicSurvey[is.na(FloristicSurvey)] <- 0
mydata<-t(data.frame(FloristicSurvey[,1],FloristicSurvey[,4:55]))
my.names <- mydata[1,]
colnames(mydata) <- my.names
mydata <- mydata[-1,]
d <- dist(t(mydata), method = "euclidean") # distance matrix
fit <- hclust(d, method = "ward.D")
plot(fit)

#do they cluster by GM coverage?
FloristicSurvey[is.na(FloristicSurvey)] <- 0
#this will just make the dendogram easier to see
FloristicSurvey$GM_Coverage_category<-cut(FloristicSurvey$GM_Coverage, c(-Inf,5,30,50,70,95,100), labels = c("N", "L", "ML", "MH", "H", "F"))
mydata<-t(data.frame(FloristicSurvey[,56],FloristicSurvey[,4:55]))
my.names <- mydata[1,]
colnames(mydata) <- my.names
mydata <- mydata[-1,]
d <- dist(t(mydata), method = "euclidean") # distance matrix
fit <- hclust(d, method = "ward.D")
plot(fit) 
FloristicSurvey$GM_Coverage_category<-cut(FloristicSurvey$GM_Coverage, c(-Inf,5,30,50,70,95,100), labels = c("None", "Low", "Med_Low", "Med_High", "High", "Overtaken"))

### NMDS 
```{r}
# https://jonlefcheck.net/2012/10/24/nmds-tutorial-in-r/
# https://oliviarata.wordpress.com/2014/04/17/ordinations-in-ggplot2/
#Make a matrix with no row or column equal to 0 (do not enclude the env variable (GM COVERAGE))
M <- as.matrix(FloristicSurvey[1:198,4:55])
M[is.na(M)] <- 0
rownames(M) <- FloristicSurvey$Population
which( colSums(M)==0 )
which( rowSums(M)==0 )
#Now that you know which column must be taken out, redo the matrix but with the last column, then remove the rows and columns =0
#I did this because the previous method doesnt work with the characters in the GM_COVERAGE column.
M <- as.matrix(FloristicSurvey[1:198,4:56])
M[is.na(M)] <- 0
rownames(M) <- FloristicSurvey$Population
M<-M[,-45]
M<-M[-130,]
M<-M[-133,]
M<-M[-132,]
# Now I'll make the env column a vector on its own, but the previous colde now allows it to be in the same order as the matrix im using for distances.
GM_coverage_df <- data.frame(M[,52])
M<-M[,-52]
class(M) <- "numeric"
# with vegdist from Bray to seroeson: add binary = T 
dist_FloristicSurvey <- vegdist(M, method = "bray", binary = T)
#The metaMDS analysis could have done the distance matrix internally but i would rather control it since i have presence/abscence
meta.nmds.FloristicSurvey <- metaMDS(dist_FloristicSurvey)
# Non convergence with only 20 tries. lets increase that then
meta.nmds.FloristicSurvey2D <- metaMDS(dist_FloristicSurvey, k=2, trymax = 1000)
str(meta.nmds.FloristicSurvey2D)
stressplot(meta.nmds.FloristicSurvey2D)
# Im getting a little above a stress of 0.2 (0.207), which isnt ideal, ill try it again with 3 dimension
# A good rule of thumb: stress < 0.05 provides an excellent representation in reduced dimensions, < 0.1 is great, < 0.2 is good/ok, and stress < 0.3 provides a poor representation.
meta.nmds.FloristicSurvey3D <- metaMDS(dist_FloristicSurvey, k=3, trymax = 1000)
str(meta.nmds.FloristicSurvey3D)
# stress is lower! 0.146
stressplot(meta.nmds.FloristicSurvey3D)

@@

## finally plotting.
mult <- 1 #multiplier for the arrows and text for envfit below. You can change this and then rerun the plot command.
FloristicSurvey.nmds.gg1 <- ggplot(data = FloristicSurvey.NMDS.data, aes(y = NMDS2, x = NMDS1))+ #sets up the plot. brackets around the entire thing to make it draw automatically
    geom_path(data = df_ell.FSurvey.GM_coverage, aes(x = NMDS1, y = NMDS2, group = df_ell.FSurvey.GM_coverage$GM_coverage, alpha=df_ell.FSurvey.GM_coverage$GM_coverage))+ #this is the ellipse, seperate ones by Site. If you didn't change the "alpha" (the shade) then you need to keep the "group 
    scale_alpha_manual(guide = FALSE,values=c(0.3, 0.5, 0.6, 0.7, 0.8, 0.9))+ #sets the shade for the ellipse
    geom_point( aes(shape = FloristicSurvey.NMDS.data$GM_Coverage), size = 1) + #puts the site points in from the ordination, shape determined by site, size refers to size of point
    # geom_text(data=spps2, aes(x=spps2$NMDS1, y=spps2$NMDS2, label=species), size = 3.3, hjust=1.1)+ #labelling the species. hjust used to shift them slightly from their points
    #annotate("text",x = NMDS.mean$NMDS1,y = NMDS.mean$NMDS2,label=NMDS.mean$group) + #labels for the centroids - I haven't used this since we have a legend. but you could also dithc the legend, but plot will get v messy
    #geom_segment(data = env.scores.FloristicSurvey,
    #             aes(x = 0, xend = mult*env.scores.FloristicSurvey$NMDS1, y = 0, yend = mult*env.scores.FloristicSurvey$NMDS2),
    #             arrow = arrow(length = unit(0.25, "cm")), colour = "grey") + #arrows for envfit.  doubled the length for similarity to the plot() function. NB check ?envfit regarding arrow length if not familiar with lengths
    #geom_text(data = env.scores.FloristicSurvey, #labels the environmental variable arrows * "mult" as for the arrows
    #          aes(x = mult*env.scores.FloristicSurvey$NMDS1, y = mult*env.scores.FloristicSurvey$NMDS2, label=env.variables),
    #          size = 5,
    #          hjust = -0.5)+
    scale_shape_manual(values = c(1,8,19,5,6,7))+ #sets the shape of the plot points instead of using whatever ggplot2 automatically provides
    coord_cartesian(xlim = c(-1,1.5))+  ## NB this changes the visible area of the plot only (this is a good thing, apparently). Can also specify ylim. Here in case you want to set xaxis manually.
    theme_simple()
```

@@

```{r}
Soil_characteristics$Location<-as.factor(Soil_characteristics$Location)
GLmodel_soil <- glm(Location ~Agg_stability*C_percent*N_percent,family=binomial(link='logit'),data=Soil_characteristics)
summary(GLmodel_soil)

@@

p <- ggplot(Soil_characteristics, aes(Location,pH,fill=Location))
p + geom_boxplot() +
  theme_simple()
```
# Histology data Analysis

>>>>

```{r}
row.has.na <- apply(PlantLength, 1, function(x){any(is.na(x))})
sum(row.has.na) 
PlantLength.filtered <- PlantLength[!row.has.na,]# Removed all the rows with NAs

car::qqp(PlantLength$Length_plant, "norm")
shapiro.test(PlantLength$Length_plant)

par(mfrow = c(1,1))
descdist(PlantLength.filtered$Length_plant, discrete = FALSE, boot = 1000)
f1g <- fitdist(PlantLength.filtered$Length_plant, "lnorm") 
f2g <- fitdist(PlantLength.filtered$Length_plant, "norm") 
f3g <- fitdist(PlantLength.filtered$Length_plant, "gamma")
f4g <- fitdist(PlantLength.filtered$Length_plant, "weibull")
f1g$aic
f2g$aic
f3g$aic
f4g$aic

mod1<-lm(Length_plant~Location*Species, data=PlantLength.filtered)

bc<-boxcox(mod1, lambda = seq(-2, 2, 0.1))
lambda <- bc$x[which.max(bc$y)]
PlantLength.filtered <- cbind(PlantLength.filtered, ((PlantLength.filtered$Length_plant^lambda)-1)/lambda)
names(PlantLength.filtered)[5] <- "Yprime"
mod1BC <- lm(Yprime~Location*Species, data=PlantLength.filtered)

par(mfrow = c(2, 2))
plot(mod1BC)
# This is basically a log transformation.

# Multicolinearity?
car::vif(mod1BC)
sqrt(car::vif(mod1BC)) > 2

mod1<-lmer(log(Length_plant)~Location*Species+(1|Pop), data=PlantLength.filtered)

mod2<-lmer(log(Length_plant)~Location+(1|Pop), data=PlantLength.filtered)

mod3<-lmer(log(Length_plant)~1+(1|Pop), data=PlantLength.filtered)

AICres <- model.sel(mod1,mod2,mod3)
r.squaredLR(mod1)
summary(mod1)
confint(mod1)
```

@@

### PCA
```{r}
colnames(FloristicSurvey)[1] <- "Population" # population loads in weirdly, I am just arranging it to not have quotes around. so odd
FloristicSurvey$Population<-as.factor(FloristicSurvey$Population) # Make the population a factor and not an interger
FloristicSurvey[is.na(FloristicSurvey)] <- 0
sum(row.has.na <- apply(FloristicSurvey, 1, function(x){any(is.na(x))}))

@@

### Linear discriminate analysis
```{r}
# Discriminant function analysis is used (general LDA format) because i have a cat. dependent and a binary independent. (it needs binary or continuous!)
# you might have to bin covers because it only works with categorical, so 0-5% none, 6-40 low etc.
FloristicSurvey$GM_Coverage_category<-cut(FloristicSurvey$GM_Coverage, c(-Inf,25,50,75,100), labels = c("None", "Med_Low", "Med_High", "Overtaken"))
table(FloristicSurvey$GM_Coverage_category)
# [0-20] None
# ]20-50] Medium low
# ]50-80] Medium high
# ]80-100] Overtaken
```

```{r}
#LDA
FloristicSurveysmall<-data.frame(FloristicSurvey[,4:56])
# Linear discriminant function analysis

@@

### Cluster analysis
```{r}
# Ward Hierarchical Clustering with Bootstrapped p values
#do they cluster by population?
FloristicSurvey[is.na(FloristicSurvey)] <- 0
mydata<-t(data.frame(FloristicSurvey[,1],FloristicSurvey[,4:55]))
my.names <- mydata[1,]
colnames(mydata) <- my.names
mydata <- mydata[-1,]
d <- dist(t(mydata), method = "euclidean") # distance matrix
fit <- hclust(d, method = "ward.D")
fitd <- as.dendrogram(fit)

par(mfrow=c(3,2))
plot(fit,cex=0.4,main="Clustering with populations")
plot(cut(fitd, h=10)$upper, main="Upper tree of cut at h=10")
plot(cut(fitd, h=10)$lower[[1]], main="First branch of lower tree with cut at h=10")
plot(cut(fitd, h=10)$lower[[2]], main="Second branch of lower tree with cut at h=10")
plot(cut(fitd, h=10)$lower[[3]], main="Third branch of lower tree with cut at h=10")
plot(cut(fitd, h=10)$lower[[4]], main="Fourth branch of lower tree with cut at h=10")

#do they cluster by GM coverage?
FloristicSurvey[is.na(FloristicSurvey)] <- 0
#this will just make the dendogram easier to see
FloristicSurvey$GM_Coverage_category<-cut(FloristicSurvey$GM_Coverage, c(-Inf,25,50,75,100), labels = c("None", "ML", "MH", "High"))
mydata<-t(data.frame(FloristicSurvey[,56],FloristicSurvey[,4:55]))
my.names <- mydata[1,]
colnames(mydata) <- my.names
mydata <- mydata[-1,]
d <- dist(t(mydata), method = "euclidean") # distance matrix
fit <- hclust(d, method = "ward.D")
fitd <- as.dendrogram(fit)
FloristicSurvey$GM_Coverage_category<-cut(FloristicSurvey$GM_Coverage, c(-Inf,5,30,50,70,95,100), labels = c("None", "Low", "Med_Low", "Med_High", "High", "Overtaken"))

par(mfrow=c(3,2))
plot(fit,cex=0.4,main="Clustering with populations")
plot(cut(fitd, h=10)$upper, main="Upper tree of cut at h=10")
plot(cut(fitd, h=10)$lower[[1]], main="First branch of lower tree with cut at h=10")
plot(cut(fitd, h=10)$lower[[2]], main="Second branch of lower tree with cut at h=10")
plot(cut(fitd, h=10)$lower[[3]], main="Third branch of lower tree with cut at h=10")
plot(cut(fitd, h=10)$lower[[4]], main="Fourth branch of lower tree with cut at h=10")
```

### NMDS 
```{r}
# https://jonlefcheck.net/2012/10/24/nmds-tutorial-in-r/
# https://oliviarata.wordpress.com/2014/04/17/ordinations-in-ggplot2/
#Make a matrix with no row or column equal to 0 (do not enclude the env variable (GM COVERAGE))
M <- as.matrix(FloristicSurvey[1:198,4:55])
M[is.na(M)] <- 0
rownames(M) <- FloristicSurvey$Population
which( colSums(M)==0 )
which( rowSums(M)==0 )
#Now that you know which column must be taken out, redo the matrix but with the last column, then remove the rows and columns =0
#I did this because the previous method doesnt work with the characters in the GM_COVERAGE column.
M <- as.matrix(FloristicSurvey[1:198,4:56])
M[is.na(M)] <- 0
rownames(M) <- FloristicSurvey$Population
M<-M[,-45]
M<-M[-130,]
M<-M[-133,]
M<-M[-132,]
# Now I'll make the env column a vector on its own, but the previous colde now allows it to be in the same order as the matrix im using for distances.
GM_coverage_df <- data.frame(M[,52])
M<-M[,-52]
class(M) <- "numeric"
# with vegdist from Bray to seroeson: add binary = T 
dist_FloristicSurvey <- vegdist(M, method = "bray", binary = T)
#The metaMDS analysis could have done the distance matrix internally but i would rather control it since i have presence/abscence

meta.nmds.FloristicSurvey2D <- metaMDS(dist_FloristicSurvey, k=2, trymax = 1000)

str(meta.nmds.FloristicSurvey2D)
stressplot(meta.nmds.FloristicSurvey2D)

@@

## finally plotting.
mult <- 1 #multiplier for the arrows and text for envfit below. You can change this and then rerun the plot command.
FloristicSurvey.nmds.gg1 <- ggplot(data = FloristicSurvey.NMDS.data, aes(y = NMDS2, x = NMDS1))+ 
    geom_path(data = df_ell.FSurvey.GM_coverage, aes(x = NMDS1, y = NMDS2, group = df_ell.FSurvey.GM_coverage$GM_coverage, alpha=df_ell.FSurvey.GM_coverage$GM_coverage))+ 
    scale_alpha_manual(guide = FALSE,values=c(0.3, 0.5, 0.6, 0.7))+ 
    geom_point( aes(shape = FloristicSurvey.NMDS.data$GM_Coverage), size = 1)+ 
    annotate("text",x = NMDS.mean.FloristicSurvey$NMDS1,y = NMDS.mean.FloristicSurvey$NMDS2,label=NMDS.mean.FloristicSurvey$group)+ 
    geom_segment(data = env.scores.FloristicSurvey,aes(x = 0, xend = mult*env.scores.FloristicSurvey$NMDS1, y = 0, yend = mult*env.scores.FloristicSurvey$NMDS2),arrow = arrow(length = unit(0.25, "cm")), colour = "grey")+ 
    geom_text(data = env.scores.FloristicSurvey,aes(x = mult*env.scores.FloristicSurvey$NMDS1, y = mult*env.scores.FloristicSurvey$NMDS2, label=env.variables),size = 5, hjust = -0.5)+ 
    scale_shape_manual(values = c(1,8,19,5))+ 
    coord_cartesian(xlim = c(-1,1.5)) 
```

@@

```{r, Soil Characteristics}
Soil_characteristics$Location<-as.factor(Soil_characteristics$Location)
GLmodel_soil <- glm(Location ~Agg_stability*C_percent*N_percent,family=binomial(link='logit'),data=Soil_characteristics)
summary(GLmodel_soil)

@@

par(mfrow=c(2,2))
ggplot(Soil_characteristics, aes(Location,Agg_stability,fill=Location))+ geom_boxplot() +
  theme_simple()
ggplot(Soil_characteristics, aes(Location,pH,fill=Location))+ geom_boxplot() +
  theme_simple()
ggplot(Soil_characteristics, aes(Location,C_percent,fill=Location))+ geom_boxplot() +
  theme_simple()
ggplot(Soil_characteristics, aes(Location,N_percent,fill=Location))+ geom_boxplot() +
  theme_simple()

```
# Histology data Analysis
